Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-08 aistats.tex
Add references for technical background

@@

\subsection{Euclide's World}
As another reference point, if we make restrictive assumptions on the log-partition function, we can ensure our objective will behave at least as nicely as a quadratic.
\paragraph{If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$, which kinda adresses \citet{bubeck2015entropic}), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    &\leq L \norm{\mu^* - \mu} + \norm{\natp} \norm{\mu^* - \mu} \\
    &\leq 2L \norm{\mu^* - \mu}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz.
\paragraph{If $\conj$ is $L$-smooth} (e.g. $\logpart$ is $L^{-1}$-strongly convex), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    \leq \frac{L}{2} \norm{\mu^* - \mu}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic, and we get the same rate as a Gaussian.

@@

\clearpage
\section{Group Meeting}
\begin{enumerate}
	\item new plots : mirror map, and gaussian trajectories. Add trajectories to paper, with background level set, and new legend : primal/dual expectation
	\item the MLE is biased according to our definition. See plots. Does this definition of bias make sense ? An unbiased estimator's primal mean would be equal to the optimum, vs the dual mean.  Remark : the MLE is biased for standard parameters.
	\item exciting pap er\citet{kakade2010learning}
	\begin{enumerate}
		\item assumption in the spirit of infinite order self-concordance only at the optimum
		\item After a burn-in phase, the loss behaves quadratically. No precise description of the burn-in length.
		\item \TODO{it is worth writing down until the end what we are obtaining with self-concordance.}
		\item idea 1 : try to express some of their results with the mean parameter instead.
		\item idea 2 : using their assumption on the mirror map, perform a Newton kind of analysis for SMD. (exploiting the link MAP -> optimization). How do we prove the linear convergence of Newton to the quadratic regime ? That is the problem at stake here.
	\end{enumerate}
	\item Intro : positioning between optimization and statistics. Start with stats. Convince the reader that this is an essential problem, with asymptotic rate. but finite is elusive. Inspired by recent work (lepriol2021 , bach2013) we want to use optimization to give a rate. We fail, but we reveal a hole in current analysis.  Finite rate would allow us to quantify the importance of the prior, and also help us understand behavior for small numbers. 
	\item Intro : 1 clear question and sharper contributions, with references to sections, as a memo for the reviewer writing his review. "investigate bias-variance..." "we show all these analysis suffer from this problem."
	\item Organization : add optim parallel in insigths section, and then sprinkle references until the optim section.
	\item Optim : find members of the exponential family where the rate of Dragomir et al. holds.
	\item Technical Background : be precise and add reference for the steep exponential family of Legendre type which guarantees that our sets are open, and have a solution in the middle.
\end{enumerate}

>>>>

\subsection{MAP as Stochastic Mirror Descent}
\TODO{s
}

@@

\subsection{Euclidean Case}
As another reference point, let us consider the case $\logpart = \half \norm{\natp}_2^2$, which corresponds to finding the mean of a Gaussian with variance $1$ ($\cX=\real, \nu(dx) = \exp(\half[-x^2])dx, T(X)=X$).

If we make restrictive assumptions on the log-partition function, we can ensure our objective will behave at least as nicely as a quadratic.
\paragraph{If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$, which kinda adresses \citet{bubeck2015entropic}), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    &\leq L \norm{\mu^* - \mu} + \norm{\natp} \norm{\mu^* - \mu} \\
    &\leq 2L \norm{\mu^* - \mu}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz.
\paragraph{If $\conj$ is $L$-smooth} (e.g. $\logpart$ is $L^{-1}$-strongly convex), then
\begin{align}
    \bregmanconj(\mu^* ; \mu) 
    \leq \frac{L}{2} \norm{\mu^* - \mu}^2
\end{align}
so $\bregmanconj$ is upper bounded by a quadratic, and we get the same rate as a Gaussian.

@@

\clearpage
\section{Group Meeting}
\begin{enumerate}
	\item new plots : mirror map, and gaussian trajectories. Add trajectories to paper, with background level set, and new legend : primal/dual expectation
	\item the MLE is biased according to our definition. See plots. Does this definition of bias make sense ? An unbiased estimator's primal mean would be equal to the optimum, vs the dual mean.  Remark : the MLE is biased for standard parameters.
	\item exciting pap er\citet{kakade2010learning}
	\begin{enumerate}
		\item assumption in the spirit of infinite order self-concordance only at the optimum
		\item After a burn-in phase, the loss behaves quadratically. No precise description of the burn-in length.
		\item \TODO{it is worth writing down until the end what we are obtaining with self-concordance.}
		\item idea 1 : try to express some of their results with the mean parameter instead.
		\item idea 2 : using their assumption on the mirror map, perform a Newton kind of analysis for SMD. (exploiting the link MAP -> optimization). How do we prove the linear convergence of Newton to the quadratic regime ? That is the problem at stake here.
	\end{enumerate}
	\item Intro : positioning between optimization and statistics. Start with stats. Convince the reader that this is an essential problem, with asymptotic rate. but finite is elusive. Inspired by recent work (lepriol2021 , bach2013) we want to use optimization to give a rate. We fail, but we reveal a hole in current analysis.  Finite rate would allow us to quantify the importance of the prior, and also help us understand behavior for small numbers. 
	\item Intro : 1 clear question and sharper contributions, with references to sections, as a memo for the reviewer writing his review. "investigate bias-variance..." "we show all these analysis suffer from this problem."
	\item Organization : add optim parallel in insigths section, and then sprinkle references until the optim section.
	\item Optim : find members of the exponential family where the rate of Dragomir et al. holds.
	\item Technical Background : be precise and add reference for the steep exponential family of Legendre type which guarantees that our sets are open, and have a solution in the middle. References: \begin{itemize}
		\item \href{https://www.jstor.org/stable/4616462?seq=1#metadata_info_tab_contents}{Existence of Maximum Likelihood Estimates for Multi-Dimensional Exponential Families}, 
		\item The singly truncated normal distribution: A non-steep exponential family, 
		\item Statistical exponential families: A digest with flash cards,
		\item Information and Exponential Families In Statistical Theory, 
		\item Fundamentals of Statistical Exponential Families with Applications in Statistical Decision Theory.
	\end{itemize}

\end{enumerate}
